---
title: "Homework 5 Answers"
author: "Xuan Beatty"
date: "2025-12-01"
output: pdf_document
---

```{r setup, include=FALSE}

knitr::opts_chunk$set(
  echo = TRUE,
  results = 'markup',
  warning = FALSE,
  message = FALSE,
  error = FALSE)

```

## Libraries needed

```{r}

library(dplyr)
library(tidyr)
library(broom)
library(purrr)
library(ggplot2)
library(stringr)

```


When you click the **Knit** button a document will be generated that includes both content as well as the output of any embedded R code chunks within the document. You can embed an R code chunk like this:

```{r reading in the data}

setwd("~/ERHS 535/Homework5_XB/data")

homicides <- read.csv("homicide-data.csv")


# Creating a new column combining city and state
homicides <- homicides %>%
  mutate(city_name = paste(city, state, sep = ", "))

```

## Including Plots

You can also embed plots, for example:

```{r pressure, echo=FALSE}
plot(pressure)
```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
